\bf Title

### 2024-11-08 article/paper_padic_spec.tex
commit by Simone
modif thm 4.1

@@

  Let $M \in \allmat_{d,d}(K)$ with characteristic polynomial $\chi_M = T^d + \sum_{i=0}^{d-1} a_i T^i$ for $a_0,...,a_{d-1} \in K$.
  Firstly, suppose   $M \not \in \allmat_{d,d}^+(K)$ {\it i.e.} $\chi_M$ have a root $\rho$ of negative valuation.
  It follows that $\val \chi_M (\rho) = \val (\rho^d+ \sum_{i=0}^{d-1} a_i \rho^i ) = + \infty$ and $\val \rho^d = d \val \rho <0$.
  The properties of the non-archimedean inequality followed by $\val $ implie that $\val ( \sum_{i=0}^{d-1} a_i \rho^i) = \val \rho^d$ and $\val ( \sum_{i=0}^{d-1} a_i \rho^i) \ge \min_{i=0,..., d-1} \val \left(a_i \rho^i\right)$. Therefore, there exists $i \in \left\{0,...,d-1\right\}$ such that $\val( a_i \rho^i) \le \val\rho^d$ {\it i.e.} such that $\val (a_i) \le (d-i) \val \rho <0$. That is all the results we need to prove the result.

  We have proven the converse sense of the equivalence. To prove the direct sense a brief introduction to Newton's polygon will be needed, we will only state the results we use in this proof and will redirect to \cite{gouvea_p-adic_2003} to further explanations.
  The Newton polygon $\mathcal{P}$ of a polynomial $P = \sum_{n}^{i=0} a_i T^i \in K[T]$ is defined as the inferior convex hull of the set $\{(i, \val a_i ) | i\in \{0,...,d\} \}$ {\it i.e.} the largest convex function $\mathcal{P} : [0,n] \to \R$ such that $\mathcal{P}(i) \leq \val a_i$ for $i=1,...,n$. 
  This function is piecewise affine and its slopes are exactly the opposites of the valuations of the roots of $P$. These are all the results we will need for this proof.

  Suppose $\chi_M \not\in \OK[T]$. There exists $j \in \{0,...,d-1\}$ such that $\val (a_j)<0$. The polynomial $\chi_M$ being monic the point with abscissa $n$ of $\chi_M$'s Newton's Polygon is $(n,0)$. 
  Therefore, the point $(j, \val a_j)$ having negative ordinate implies that at least one of the slopes of $\chi_M$'s Newton's Polygon is positive so that $\chi_M$ has a root with negative valuation.
\end{proof}

>>>>

Recall that the {\it Newton polygon} $\mathcal{P}$ of a polynomial $P = \sum_{i=0}^{n} a_i T^i \in K[T]$ is
defined as the inferior convex hull of the set $\{(i, \val a_i) \,|\, i\in \{0,...,d\} \}$, {\it i.e.} the
graph of the greatest convex function $\mathcal{P} : [0,n] \to \R$ such that $\mathcal{P}(i) \leq \val a_i$
for $i=1,...,n$. This function is piecewise affine and its slopes are exactly the opposites of the valuations
of the roots of $P$.
  %%%These are all the results we will need for this proof.

@@

  Let $M \in \allmat_{d,d}(K)$ with characteristic polynomial $\chi_M = T^d + \sum_{i=0}^{d-1} a_i T^i$ for $a_0,...,a_{d-1} \in K$.
  Firstly, suppose $M \not \in \allmat_{d,d}^+(K)$, {\it i.e.} $\chi_M$ has a root $\rho$ with $\val\,\rho<0$.
  It follows that
  $$
  \val \chi_M (\rho) = \val \left(\rho^d+ \sum_{i=0}^{d-1} a_i \rho^i\right) = + \infty
  $$ and $\val \rho^d = d \val \rho <0$.
  The properties of the non-archimedean inequality and of the valuation imply that
  $$
  \min_{0 \leq i \leq d-1} \val \left(a_i \rho^i\right) \leq \val\left(\sum_{i=0}^{d-1} a_i \rho^i\right) = \val \rho^d
  $$
  %%%$\val ( \sum_{i=0}^{d-1} a_i \rho^i) = \val \rho^d$ and $\val ( \sum_{i=0}^{d-1} a_i \rho^i) \ge \min_{i=0,..., d-1} \val \left(a_i \rho^i\right)$.
  Therefore, there exists $j \in \left\{0,...,d-1\right\}$ such that $\val(a_j \rho^j) \le \val\rho^d$,
  {\it i.e.} such that $\val a_j \le (d-j) \val \rho <0$. This shows that $\chi_M \not\in \OK[T]$.

  %%We have proven the converse sense of the equivalence.
  %%To prove the direct sense a brief introduction to Newton's polygon will be needed, we will only state the results we use in this proof and will redirect to \cite{gouvea_p-adic_2003} to further explanations.

  Suppose now $\chi_M \not\in \OK[T]$. There exists $j \in \{0,...,d-1\}$ such that $\val a_j <0$. The polynomial $\chi_M$ being monic, the point with abscissa $n$ of the Newton polygon of $\chi_M$ is $(n,0)$. 
  Therefore, the point $(j, \val a_j)$ having negative ordinate implies that at least one of the slopes of
  the Newton polygon of $\chi_M$ is positive, so that $\chi_M$ has a root with negative valuation.
\end{proof}
